single_elimination

### 2025-11-07 group_stage.qmd
Updates

```{r}
simulate_group_stage_tournament <- function(num_teams, distribution, group_length = 4, ties = TRUE, rounds = 1, seeded=F) {

@@

  # Create team strengths based on distribution
  if (distribution == "Normal"){
    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
    team_strengths <- data.frame(
      true_rank = as.numeric(teams),
      theta = strengths,
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(team_strengths, true_rank)
  }
  else if (distribution == "Same"){
    strengths <- sapply(num_teams, function(n) {rep(0, n)})
    team_strengths <- data.frame(
      true_rank = as.numeric(teams),
      theta = strengths,
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(team_strengths, true_rank)
  }
  else if (distribution == "Uniform"){
    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1), 0, sqrt(12)) })
    team_strengths <- data.frame(
      true_rank = as.numeric(teams),
      theta = strengths,
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams)),
      rank_hat = rep(NA, length(teams))
    )
    df <- arrange(team_strengths, true_rank)
  }
  else if (distribution == "Manual") {
    cat("Please enter the strengths for each team in descending rank order:\n")
    strengths <- numeric(num_teams)
    for (i in num_teams:1) {
      repeat {
        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
        if (!is.na(input)) {
          strengths[i] <- input
          break
        } else {
          cat("Invalid input. Please enter a numeric value.\n")
        }
      }
    }
    team_strengths <- data.frame(
      true_rank = as.numeric(teams),
      theta = strengths,
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(team_strengths, true_rank)
  }
  else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

@@

  group_winners <- final_df %>%
    filter(group_rank == 1) %>%
    arrange(groups)

  group_seconds <- final_df %>%
    filter(group_rank == 2) %>%
    arrange(groups)

@@

  knockout_results <- simulate_knockout(df = final_df, matchup_vector = matchup_vector, series = 1, third_place = TRUE)

@@

  combined_df <- bind_rows(
  final_df %>% filter(rank_hat >= num_groups),   # group-stage only teams
  knockout_results %>% filter(!is.na(rank_hat))  # knockout-stage teams
) %>%
  arrange(true_rank) %>%
    mutate(rank_hat = rank(rank_hat, ties.method = "random"))

@@

# Example usage:
# Single simulation
result <- simulate_group_stage_tournament(num_teams = 8, distribution = "Normal", group_length = 4, ties=F)

>>>>

```{r}
simulate_group_stage_tournament <- function(num_teams, distribution, group_length = 4, ties = TRUE, rounds = 1, third_place = T, theta_hat = NULL) {

@@

  # Create team strengths based on distribution
  if (distribution == "Normal"){
    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
    team_strengths <- data.frame(
      true_rank = as.numeric(teams),
      theta = strengths,
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(team_strengths, true_rank)
  }
  else if (distribution == "Same"){
    strengths <- sapply(num_teams, function(n) {rep(0, n)})
    team_strengths <- data.frame(
      true_rank = as.numeric(teams),
      theta = strengths,
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(team_strengths, true_rank)
  }
  else if (distribution == "Uniform"){
    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1), 0, sqrt(12)) })
    team_strengths <- data.frame(
      true_rank = as.numeric(teams),
      theta = strengths,
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams)),
      rank_hat = rep(NA, length(teams))
    )
    df <- arrange(team_strengths, true_rank)
  }
  else if (distribution == "Manual") {
    strengths <- theta_hat
    manual_strengths <- data.frame(
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- manual_strengths %>% arrange(true_strength) %>%
      mutate(true_rank = as.numeric(teams)) %>%
      arrange(true_strength, true_rank)
  }
  else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

@@

  # Get top finishers from each group (group_rank <= 1.5 covers both rank 1 and tied 1.5)
  group_winners <- final_df %>%
    group_by(groups) %>%
    filter(group_rank == min(group_rank)) %>%  # Get best finisher(s)
    slice(1) %>%  # If still tied, take first one
    ungroup() %>%
    arrange(groups)

  # Get second place finishers (next best after winners)
  group_seconds <- final_df %>%
    group_by(groups) %>%
    arrange(group_rank, true_rank) %>%  # Sort by rank, then true_rank as tiebreaker
    slice(2) %>%  # Take second team
    ungroup() %>%
    arrange(groups)

@@

  knockout_results <- simulate_knockout(df = final_df, matchup_vector = matchup_vector, series = 1, third_place = third_place)

@@

  combined_df <- bind_rows(
  final_df %>% filter(rank_hat >= num_groups),   # group-stage only teams
  knockout_results %>% filter(!is.na(rank_hat))  # knockout-stage teams
) %>%
  arrange(true_rank)
  if (ties == F){
    final_df <- final_df %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
  } else{
    final_df <- final_df %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
  }

@@

# Example usage:
# Single simulation
result <- simulate_group_stage_tournament(num_teams = 16, distribution = "Same", group_length = 4, ties=T)

### 2025-11-07 mutual_information_curves.qmd
Updates

weighted_kendall(c(2,1,3,4), k=4, salient_weights = c(3,1,1,0))
weighted_kendall(c(1.5,3,1.5), k=2)
weighted_kendall(c(1,3,4,2), k=3)
weighted_kendall(c(1,1,4,1), k=3)

@@

Triple Elimination

@@

Round Robin - with k rounds (1,2,3,5,7)

Stepladder - 8 vs 7, winner plays 6, etc.

@@

Staged Round Robin/Jacobian Ladder - top 2 in each group form new group, bottom 2 new group (each new group is a new round robin)

>>>>

weighted_kendall(rhat=c(2,1,3,4), k=4)
weighted_kendall(c(1.5,3,1.5), k=2)
weighted_kendall(c(1,3,4,2), k=3)
weighted_kendall(c(1,1,4,1), k=3)

@@

~~Triple Elimination~~

@@

Round Robin - with m rounds (1,2,4)

Stepladder- 8 vs 7, winner plays 6, etc.

Waterfall - stepladder round robin

@@

Staged Round Robin - top 2 in each group form new group, bottom 2 new group (each new group is a new round robin)

### 2025-11-07 single_elimination_shiny.qmd
Updates

```{r}
# Simulate the single-elimination tournament
simulate_single_elimination_tournament <- function(num_teams, distribution, seeding_structure = NULL, ties = T, series = 1, third_place = T, reseeded = F) {

  if (num_teams <= 3) {
    stop("Number of Teams must be greater than 3.")
  }
  teams <- paste(num_teams:1)
  if (log2(num_teams) %% 1 != 0) {
    power_of_2 <- 2^ceiling(log2(num_teams))
    teams <- c(teams, rep(NA, power_of_2 - num_teams))
  }

  if (distribution == "Normal"){
    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Same"){
    strengths <- sapply(num_teams, function(n) {rep(0,n)})
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Uniform"){
    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
    unif_strength <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams)),
      rank_hat = rep(NA,length(teams))
    )
    df <- arrange(unif_strength, true_rank)
  }
  else if (distribution == "Manual") {
    cat("Please enter the strengths for each team in descending rank order:\n")
    strengths <- numeric(num_teams)
    for (i in num_teams:1) {
      repeat {
        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
        if (!is.na(input)) {
          strengths[i] <- input
          break
        } else {
          cat("Invalid input. Please enter a numeric value.\n")
        }
      }
    }
    manual_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(manual_strengths, true_rank)
  }
  else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

@@

      # RESEEDING LOGIC: If reseeded is TRUE, sort teams by their true rank before each round
      if (reseeded && length(teams) > 1) {
        # Remove NA teams for sorting
        valid_teams <- teams[!is.na(teams)]
        na_teams <- teams[is.na(teams)]

@@

        # Create reseeded matchups: 1 vs worst, 2nd vs 2nd worst, etc.
        n_valid <- length(sorted_teams)
        if (n_valid > 1) {
          reseeded_teams <- c()
          for (j in 1:(n_valid/2)) {
            reseeded_teams <- c(reseeded_teams, sorted_teams[j], sorted_teams[n_valid - j + 1])
          }
          # Add back any NA teams (byes)
          teams <- c(reseeded_teams, na_teams)
        }
      }

>>>>

```{r}
# Simulate the single-elimination tournament
simulate_single_elimination_tournament <- function(num_teams, distribution, seeding_structure = NULL, ties = T, series = 1, third_place = T, reseeded = F, theta_hat = NULL) {

  if (num_teams <= 3) {
    stop("Number of Teams must be greater than 3.")
  }

  teams <- paste(num_teams:1)

  if (log2(num_teams) %% 1 != 0) {
    power_of_2 <- 2^ceiling(log2(num_teams))
    teams <- c(teams, rep(NA, power_of_2 - num_teams))
  }

  if (distribution == "Normal"){
    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Same"){
    strengths <- sapply(num_teams, function(n) {rep(0,n)})
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Uniform"){
    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
    unif_strength <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams)),
      rank_hat = rep(NA,length(teams))
    )
    df <- arrange(unif_strength, true_rank)
  }
  else if (distribution == "Manual") {
    strengths <- theta_hat
    manual_strengths <- data.frame(
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- manual_strengths %>% arrange(true_strength) %>%
      mutate(true_rank = as.numeric(teams)) %>%
      arrange(true_strength, true_rank)
  }
  else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

@@

      # RESEEDING LOGIC: If reseeded is TRUE, sort teams by their true rank before each round
      if (reseeded && length(teams) > 1) {
        # Remove NA teams for sorting
        valid_teams <- teams[!is.na(teams)]
        num_byes <- sum(is.na(teams))

@@

        # Create reseeded matchups with byes going to top seeds
        n_valid <- length(sorted_teams)

        if (n_valid > 1) {
          reseeded_teams <- c()

          # If there are byes, give them to the top seeds
          if (num_byes > 0) {
            # Top seeds get byes
            for (j in 1:num_byes) {
              reseeded_teams <- c(reseeded_teams, sorted_teams[j], NA)
            }
            # Remaining teams play each other
            remaining_teams <- sorted_teams[(num_byes + 1):n_valid]
            n_remaining <- length(remaining_teams)
            if (n_remaining > 1) {
              for (j in 1:(n_remaining/2)) {
                reseeded_teams <- c(reseeded_teams, remaining_teams[j], remaining_teams[n_remaining - j + 1])
              }
            }
          } else {
            # No byes: standard reseeding (best vs worst)
            for (j in 1:(n_valid/2)) {
              reseeded_teams <- c(reseeded_teams, sorted_teams[j], sorted_teams[n_valid - j + 1])
            }
          }

          teams <- reseeded_teams
        }
      }
